---
title: 'VariSel : An a R package to perform variable selection in linear models'
author: "Marie"
date: "3 février 2019"
output:
  pdf_document: default
  html_document: default
header-includes:
- \usepackage{tikz}
- \usepackage{natbib}
- \usepackage{pgfplots}
- \newcommand{\Am}{\textrm{Argmin}}
- \newcommand{\bX}{\boldsymbol{X}}
- \newcommand{\bY}{\boldsymbol{Y}}
- \newcommand{\bS}{\boldsymbol{S}}
- \usetikzlibrary{arrows, decorations.markings}
- \usepackage{arydshln}
- \definecolor{outcome}{RGB}{165,77,106}
- \definecolor{darkred}{rgb}{0.65,0.15,0.25}
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE, dev = "cairo_pdf")
require(VariSel)
require(tidyverse)
require(tidymodels)
require(viridis)
require(furrr)
require(rlang)
require(knitr)
require(ggsci)
require(BlockCov)
theme_set(theme_bw() + 
            theme(strip.background = element_rect(fill = "white", color ="gray"), 
                  strip.text = element_text(color = "black"),
                  text = element_text(face="bold", family="LM Roman 10"),
                  plot.title = element_text(hjust = 0.5, size =32)))
scale_fill_continuous <- function(...) scale_fill_viridis()
scale_fill_discrete <- function(...) scale_fill_viridis(discrete = TRUE)
scale_colour_discrete <- function(...) scale_color_viridis()
scale_colour_discrete <- function(...) scale_color_viridis(discrete = TRUE)


```

# Introduction 
## Statistical modelling 

\begin{itemize}
\item \textbf{Dataset description:}

\begin{itemize}
\item $\boldsymbol{X}$:  $n\times p$ design matrix 
\item $\boldsymbol{Y}$:  $n\times q$  response matrix 
\end{itemize}

\item \textbf{Question:} Which variables influence the responses?

\item \textbf{Approach:}
\begin{itemize}
 \item Variable selection in 
$$
\boldsymbol{Y}=\boldsymbol{XB}+\boldsymbol{E},
$$
where  \begin{itemize}
\item $\boldsymbol{B}$:   $p\times q$  \textbf{sparse} coefficients matrix
\item $\boldsymbol{E}$: $n\times q$ error matrix  with $$\forall  i\in\{1,\dots, n\}, \; (E_{i,1},\dots,E_{i,q})\stackrel{iid}{\sim}~\mathcal{N}(0,\boldsymbol{\Sigma}_q)$$

\end{itemize}
\item We take the dependence into account by estimating $\boldsymbol{\Sigma}_q$. 
\end{itemize}
\end{itemize}


## Differents penalties : for different point of view

\begin{itemize}
\item \textbf{Lasso} : \textit{select variables without taking into account  potential links.}
\begin{equation*} 
	\widehat{b}_L = \Am_{b} \left\lbrace||y-\mathcal{X} b||_2^2 + \lambda ||b||_1 \right\rbrace,
\end{equation*}
\item \textbf{Group-Lasso}  : \textit{select a group of variables.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_G = \Am_{b_1, \dots, b_L} \left\lbrace||y-\sum_{1 \leq \ell \leq L}\mathcal{X}_{(\ell)} b_{(\ell)}||_2^2 + \lambda\sum_{1 \leq \ell \leq L}\sqrt{p_\ell} ||b_\ell||_2\right\rbrace,
\end{equation*}
\item \textbf{Fused-Lasso} :  \textit{influence a group of variables to have the same coefficient.}
\begin{equation*}\label{fuselasso}
	\widehat{b}_F =\Am_{b}||y-\mathcal{X} b||_2^2 + \left\lbrace\lambda_1\sum_{(i,j) \in \mathcal{G}} |b_i - b_j|+ \lambda_2 ||b||_1\right\rbrace,
\end{equation*}
\end{itemize}


This differents penalties are here in univariate. In order to use them and in order to take into account the dependance that may exist among variables we propose the following transformation : 

 $\boldsymbol{Y}\widehat{\boldsymbol{\Sigma}}_q^{-1/2}  
    =\boldsymbol{X}\boldsymbol{B}\widehat{\boldsymbol{\Sigma}}_q^{-1/2} ++ \boldsymbol{E}\widehat{\boldsymbol{\Sigma}}_q^{-1/2}$
    \begin{align*}
      {\mathcal{Y}}&=
                     vec(\boldsymbol{Y}\widehat{\boldsymbol{\Sigma}}_q^{-1/2})  
=vec(\boldsymbol{X}\boldsymbol{B}\widehat{\boldsymbol{\Sigma}}_q^{-1/2})
                     +vec(\boldsymbol{E}\widehat{\boldsymbol{\Sigma}}_q^{-1/2})\\
                   &=\textcolor{blue}{((\widehat{\boldsymbol{\Sigma}}_q^{-1/2})'\otimes \boldsymbol{X})}\textcolor{orange}{vec(\boldsymbol{B})}
                     +\textcolor{darkred}{vec(\boldsymbol{E}\widehat{\boldsymbol{\Sigma}}_q^{-1/2})}\\
                   &=\textcolor{blue}{\mathcal{X}}\textcolor{orange}{\mathcal{B}}+\textcolor{darkred}{\mathcal{E}}.
    \end{align*} 

This transformation need the estimation of the square root of the inverse of the covariance matrix $\boldsymbol{\Sigma}_q$.


# Package Installation 

```{r, eval= FALSE}
devtools::install_github("Marie-PerrotDockes/VariSel")
```


# First exemples : Iris dataset

```{r}
library(car)
iris %>% 
  head() %>% 
  kable()
```

### Associtaion between sepal characteristics and petal characteristics

The aim of this exemple is to select association between the Sepal (Length and width) and the Petal (Length and Width). 

Construction of the matrices  :

```{r}
Y <- iris %>% select( starts_with("Sepal"))
Petal_char <- iris %>% select( starts_with("Petal"))
```

Coefficient estimation using a lasso criterion: 


```{r}
mod <-  train_VariSel( Y = Y, 
                       X = Petal_char, 
                       type ="lasso_univ")
```

```{r}
plot(mod)
```

It start by selecting an association between the petal length and the sepal length and then an association between the petal lenght and the sepal widthand so on..

Let us now see what happened when we use group lasso to force the association to be between one sepal characteristics with all petal characteristics. 

```{r}
mod_group <-  train_VariSel( Y = Y, 
                       X = Petal_char, 
                       type ="group_multi_regr", 
                       sepx = "\\.")
```

The argument ~sepx = "\\."~ mean that the name of X will be seperate into a group name and a characteristics name. 
Two variables having the same group name will be selected together. Here both Petal.length and Petal.Width start with Petal hence they will be selected together. 

```{r}
plot(mod_group)
```

Let's now compare differents types grouping or not the petal characteristics and influence it or not to have the same coefficients. 

```{r}
 plan(multiprocess)
ct <- compar_type( Y = Y, X = Petal_char,  
  types = c("group_multi_regr" , "group_multi_both" ,
  "fused_multi_regr", "fused_multi_both",
  "lasso_multi", "lasso_univ" ), times = 10)
plot_ct(ct)
```

We can also propose different associations between petal characteristics and sepal charcateristics depending on the species. 


```{r}
mod_group_species <-  train_VariSel( Y = Y, 
                       regressors =  Petal_char, 
                       group = iris$Species,
                       type ="group_multi_regr")
```

```{r}
plot(mod_group_species)
```

```{r, echo=FALSE}
col <- pal_uchicago()(6)

mod_group_resp_species <-  train_VariSel( Y = Y,
                       regressors = Petal_char, 
                       group = iris$Species, 
                       type ="group_multi_both")

mod_fused_multi_species <-  train_VariSel( Y = Y,
                       regressors = Petal_char, 
                       group = iris$Species, 
                       type ="fused_multi_regr")
mod_fused_multi_species_resp <-  train_VariSel( Y = Y,
                       regressors = Petal_char, 
                       group = iris$Species, 
                       type ="fused_multi_both")


```

```{r}
compar_path(mods = list(mod_group_species,
                        mod_group_resp_species,
                        mod_fused_multi_species,
                        mod_fused_multi_species_resp))
```


```{r}
mod_lasso <-   train_VariSel( Y = Y,
                       regressors = Petal_char, 
                       group = iris$Species, 
                       type ="lasso_multi")
```
 
 
```{r}
plot(mod_lasso)
```
 

## A more serious exemple 

we will now use the dataset available and described in "A global dataset of CO2 emissions and ancillary data related to emissions for 343 cities"

The data are available in https://www.nature.com/articles/sdata2018280?WT.feed_name=subjects_environmental-social-sciences

```{r}
CO2 <-read.csv(file="Data/D_FINAL.csv", dec =",")
dim(CO2)
cc <-complete.cases(CO2[,c("Scope.1.GHG.emissions..tCO2.or.tCO2.eq.",
                           "Total.emissions..CDP...tCO2.eq.",
                           "Population..CDP.",
                           "Average.annual.temperature..CDP...degrees.Celsius.",
                           "Average.altitude..m.")])
CO2_full <- CO2[cc, ]
CO2_quant <- CO2_full %>% select_if(is.numeric)
CO2_quant_full <- CO2_quant[,complete.cases(t(CO2_quant))]
Region <- as.character(CO2_full$Region)
Region[grepl('Asia', Region)] <- "Asia"
kable(table(Region))
```


```{r}
Y <- CO2_quant_full[,c("Scope.1.GHG.emissions..tCO2.or.tCO2.eq.",
                      "Total.emissions..CDP...tCO2.eq.")] %>% 
  scale()
colnames(Y) <- c("Scope1", "Total")
CO2_quant_full_exp <-CO2_quant_full %>%
  select(-Scope.1.GHG.emissions..tCO2.or.tCO2.eq.,
    -Total.emissions..CDP...tCO2.eq.) %>% scale()
```


First model : association between in different region 


```{r}
mod_lasso <- train_VariSel(Y = Y, 
                           X = CO2_quant_full_exp,
                           type = "lasso_multi")
plot(mod_lasso)
```


You can choose to change the number of color variable (the one we can see as selected)

```{r}
plot(mod_lasso, n=20)
```


```{r}
mod_lasso_reg <- train_VariSel(Y = Y, 
                           regressors = CO2_quant_full_exp,
                           group = Region,
                           type = "lasso_multi")
plot(mod_lasso_reg)
```



```{r}
mod_group_multi_regr<- train_VariSel(Y = Y, 
                           regressors = CO2_quant_full_exp,
                           group = as.character(CO2_full$Region),
                           type = "group_multi_regr")
plot(mod_group_multi_regr, n =8)
```

we also easily have acces to coefficient : 

```{r}
kable(head(coef(mod_group_multi_regr)))
```



```{r}
mod_lasso_reg <- train_VariSel(Y = Y, 
                           regressors = CO2_quant_full_exp,
                           group = as.character(CO2_full$Region),
                           type = "group_multi_both")
plot(mod_lasso_reg, n =6)
```


```{r}
mod_lasso_reg <- train_VariSel(Y = Y, 
                           regressors = CO2_quant_full_exp,
                           group = as.character(CO2_full$Region),
                           type = "fused_multi_both")
plot(mod_lasso_reg, n =6)
```



```{r}
set.seed(123)
plan(multiprocess)
ct <- compar_type(Y = Y, 
        regressors = CO2_quant_full_exp,
        group = as.character(CO2_full$Region),
        types = c("group_multi_regr" , "group_multi_both" ,
          "fused_multi_regr", "fused_multi_both",
          "lasso_multi", "lasso_univ" ))
plot_ct(ct)

bm <- get_best_models(ct, criterion ="MSE_boot")


plot_md(bm, type = "fused_multi_both")
```


Even when influence by the model to have the same coefficient the explicative variables keep different coefficient according to the place 

```{r}

plot_md(bm)
```
 

## When there is more reponses than  individuals 


```{r}
require(MultiVarSel)
dim(MultiVarSel::prot)
kable(head(prot[, 1:10]))
```

In this case the number of reponses is the nulbver of proteines (724 for 9 individuals )

```{r}
mod_lasso_reg <- train_VariSel(Y =prot[,-1], 
                           regressors = NULL,
                           group = prot[,1],
                           type = "lasso_multi",
                           type_S12_inv = "Block")
plot(mod_lasso_reg)
kable(head(coef(mod_lasso_reg)))
```



