---
title: "VariSel: An R package to perform variable selection in multivariate linear model"
subtitle: "Or how can we preserve the wine of South of France ?"
author: "Marie Perrot-Dockès, Julien Chiquet, Laure Sansonnet"
date: "20 juin 2019"
output: 
  beamer_presentation :
    theme: "Rochester"
    colortheme: "vert"
header-includes:
- \usepackage{tikz}
- \usepackage{natbib}
- \usepackage{pgfplots}
- \newcommand{\Am}{\textrm{Argmin}}
- \newcommand{\bX}{\boldsymbol{X}}
- \newcommand{\bY}{\boldsymbol{Y}}
- \newcommand{\bS}{\boldsymbol{S}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
library(VariSel)
library(tidyverse)
```

# A simlpe vision of the immune system :
\framesubtitle{Dialogue between Th lymphocyte and dendritic-cells}
 

```{tikz, tikz-ex2, fig.cap = "DC Th dialogue", fig.ext = 'pdf', echo =FALSE}
\definecolor{vert}{RGB}{ 109,205,89}
\definecolor{bleu}{RGB}{93,139,172}
\definecolor{marron}{RGB}{172,126,93}
\definecolor{marronfonce}{RGB}{120,88,65}
\definecolor{i1}{RGB}{122,181,168}
\definecolor{i2}{RGB}{122,165,181}
\definecolor{i3}{RGB}{122,181,139}
\definecolor{dc}{RGB}{206,184,181}
\definecolor{dc2}{RGB}{123, 87, 229}
\definecolor{lt}{RGB}{181,203,206}
\tikzstyle{line} = [draw, -latex, color = marron, line width=1pt]
\tikzstyle{block} = [rectangle, draw, fill=vert!20, draw= vert!80, text = marronfonce,
text width=3.5em, text centered, rounded corners, minimum height=2em,line width=1pt]
\tikzstyle{block3} = [rectangle, draw, fill=vert!10, draw= vert!90, text = marronfonce,
text width=3em, text centered, rounded corners, minimum height=1.1em,line width=0.7pt]
\tikzstyle{inp} = [rectangle, draw,  text = marronfonce,
text width=4.4em, text centered, rounded corners, minimum height=1em,line width=1pt]
  \begin{tikzpicture}[node distance = 2cm, auto]
    % nodes
     \node(DC){\includegraphics[width= 0.24\textwidth]{dendritic-cell.png}}; 
      \node[left of = DC, node distance = 3 cm,  color = gray, xshift =-1cm](Pert){Perturbators}; 
       \node[right of = DC, node distance = 4 cm](TL){\includegraphics[width= 0.19\textwidth]{lt4-crop.pdf}}; 
       \node[right of = TL, node distance = 3 cm](L){};
        \node[below of = DC, node distance =2.2 cm,text width= 0.22\textwidth,text centered,color =dc!60!black]{Monocyte derived \;  \; Dendritic cell};
      \node[below of = TL, node distance =2.2 cm,text width= 0.22\textwidth, text centered,color =lt!60!black]{ Naive Th lymphocytes};
      \node[above of = TL, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 1.7 cm, color =lt!70!black]{ Th signals};
      \node[above of =DC, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 2.2 cm, color =dc!70!black]{ DC signals};
    
       \path [line,color = dc] ([yshift=-0.8 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.8 cm, xshift = 0.6cm]TL.west);
        \path [line,color = dc] ([yshift=-0.4 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.4 cm, xshift = 0.6cm]TL.west);
         \path [line,color = dc] ([xshift=- 0.1 cm]DC.east) --  ([xshift = 0.6cm]TL.west);
         \path [line,color = dc] ([yshift=0.4 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.4 cm, xshift = 0.6cm]TL.west);
          \path [line,color = dc] ([yshift=0.8 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.8 cm, xshift = 0.6cm]TL.west);

\path [line, color = gray](Pert) -- ([xshift =-0.6cm]DC.west);
 \path [line,color = lt]([xshift= -0.4cm, yshift = -0.7cm]TL.east) -- ([yshift = -0.7cm]L.west);
  \path [line,color = lt]([xshift= -0.4cm, yshift = -0.35cm]TL.east) -- ([yshift = -0.35cm]L.west);
  \path [line,color = lt] ([xshift= -0.4cm]TL.east) -- (L.west);
  
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.35cm]TL.east) -- ([yshift = 0.35cm]L.west);
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.7cm]TL.east) -- ([yshift = 0.7cm]L.west);
        \end{tikzpicture}

\begin{tikzpicture}[node distance = 2cm, auto]
   \node(DC){\includegraphics[width= 0.24\textwidth]{dendritic-cell2.png}}; 
      \node[left of = DC, node distance = 3 cm,  color = gray, xshift =-1cm](Pert){Perturbators}; 
       \node[right of = DC, node distance = 4 cm](TL){\includegraphics[width= 0.19\textwidth]{lt4-crop.pdf}}; 
       \node[right of = TL, node distance = 3 cm](L){};
        \node[below of = DC, node distance =2.2 cm,text width= 0.22\textwidth,text centered,color =dc2!60!black]{Blood \; \; \; Dendritic cell}; 
      \node[below of = TL, node distance =2.2 cm,text width= 0.22\textwidth, text centered,color =lt!60!black]{ Naive Th lymphocytes};
      \node[above of = TL, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 1.7 cm, color =lt!70!black]{ Th signals};
      \node[above of =DC, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 2.2 cm, color =dc2!70!black]{ DC signals};
    
       \path [line,color = dc2] ([yshift=-0.8 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.8 cm, xshift = 0.6cm]TL.west);
        \path [line,color = dc2] ([yshift=-0.4 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.4 cm, xshift = 0.6cm]TL.west);
         \path [line,color = dc2] ([xshift=- 0.1 cm]DC.east) --  ([xshift = 0.6cm]TL.west);
         \path [line,color = dc2] ([yshift=0.4 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.4 cm, xshift = 0.6cm]TL.west);
          \path [line,color = dc2] ([yshift=0.8 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.8 cm, xshift = 0.6cm]TL.west);

\path [line, color = gray](Pert) -- ([xshift =-0.6cm]DC.west);
 \path [line,color = lt]([xshift= -0.4cm, yshift = -0.7cm]TL.east) -- ([yshift = -0.7cm]L.west);
  \path [line,color = lt]([xshift= -0.4cm, yshift = -0.35cm]TL.east) -- ([yshift = -0.35cm]L.west);
  \path [line,color = lt] ([xshift= -0.4cm]TL.east) -- (L.west);
  
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.35cm]TL.east) -- ([yshift = 0.35cm]L.west);
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.7cm]TL.east) -- ([yshift = 0.7cm]L.west);

        \end{tikzpicture}

```




# Statistical modelling 

\begin{itemize}
\item \textbf{Dataset description:}

\begin{itemize}
\item $\boldsymbol{X}$:  $n\times p$ design matrix
\item $\boldsymbol{Y}$:  $n\times q$  response matrix
\end{itemize}

\item \textbf{Question:} Which variables influence the responses?

\item \textbf{Approach:}
\begin{itemize}
 \item Variable selection in 
$$
\boldsymbol{Y}=\boldsymbol{XB}+\boldsymbol{E},
$$
where  \begin{itemize}
\item $\boldsymbol{B}$:   $p\times q$  \textbf{sparse} coefficients matrix
\item $\boldsymbol{E}$: $n\times q$ error matrix  with $$\forall  i\in\{1,\dots, n\}, \; (E_{i,1},\dots,E_{i,q})\stackrel{iid}{\sim}~\mathcal{N}(0,\boldsymbol{\Sigma}_q)$$

\end{itemize}
\item We take the dependence into account by estimating $\boldsymbol{\Sigma}_q$. 
\end{itemize}
\end{itemize}


# Differents penalties : for differents point of view

\begin{itemize}
\item \textbf{Lasso} : \textit{select variables without taking into account  potential links.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_L = \Am_{b} \left\lbrace||y-\bX b||_2^2 + \lambda ||b||_1 \right\rbrace,
\end{equation*}
\item \textbf{Group-Lasso}  : \textit{select a group of variables.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_G = \Am_{b_1, \dots, b_L} \left\lbrace||y-\sum_{1 \leq l \leq L}\bX_{(l)} b_{(l)}||_2^2 + \lambda\sum_{1 \leq l \leq L}\sqrt{p_l} ||b_l||_2\right\rbrace,
\end{equation*}
\item \textbf{Fused-Lasso} \citep{xin2014efficient}   :  \textit{influence a group of variables to have the same coefficient.}
\begin{equation*}\label{fuselasso}
	\widehat{b}_F =\Am_{b}||y-\bX b||_2^2 + \left\lbrace\lambda_1\sum_{(i,j) \in \mathcal{G}} |b_i - b_j|+ \lambda_2 ||b||_1\right\rbrace,
\end{equation*}
\end{itemize}

# Concretely the use of the differents penalties in our  application

\begin{itemize}
\item \textbf{Lasso} can: 
\begin{itemize}
\item Select associations between one DC signal and one  Th signal for one DC type.
\end{itemize}
\item \textbf{Group-Lasso}  can: 
\begin{itemize}
\item Select associations between one DC signal and one  Th signal for both DC type.
\item Select associations between one DC signals and all Th signals for one DC type.
\item Select associations between one DC signals and all Th signals for both DC type.
\end{itemize}
\item \textbf{Fused-Lasso} can  : 
\begin{itemize}
\item Influence associations between one DC signal and one  Th signal to be the same for both DC type.
\item Influence associations between one DC signals and all Th signals to be the same  for one DC type.
\item Influence associations between one DC signals and all Th signals to be the same  for both DC type.
\end{itemize}
\end{itemize}


# Exemple of coefficients matrices


#  VariSel for one model type 


```{r, include=FALSE}
load("raw_data.RData")
load("data_Tmod_sspdc.RData")
DC_sign <- Xbrt2[,30:36]
T_sign <- Ybrt2[,2:5]
dc <- as.character(dc)
```

```{r, eval=FALSE}
X <-  model.matrix(~DC_sign:dc -1)
mod <-  train_VariSel( Y = T_sign, 
                       X = X, 
                       sepx = ":", 
                       type ="group_multi_marker",
                       type_S12_inv= "emp")
```

```{r}
mod <-  train_VariSel( Y = T_sign,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_marker")
```


<!-- Different parameters :  -->
<!-- \begin{itemize} -->
<!-- \item $\bY$ : Matrix containing the reponse values (here the Th signals for the different samples) -->
<!-- \item $\bX$ : optional design matrix of the model  -->
<!-- \item \textbf{type} : type of model -->
<!-- \item \textbf{regressors} : if X is not supplied the continuous values  -->
<!-- \item \textbf{group} : if the samples can be divided into groups  -->
<!-- \item \textbf{sepx} : if the columns of X can be divided into group -->
<!-- \item \textbf{type\_S12\_inv} : specified the method use to estimate $\bS$ -->
<!-- \end{itemize} -->




#  VariSel for one model type : Outcome 



# Models selection 


# Models comparison 


# Let's come back to wine!


# Conclusion

# What I have not tell you 

Come and see the vignette!

# Any questions ? 
