---
title: "VariSel: An R package to perform variable selection in the multivariate linear model"
subtitle: "Or how a Gallic village stays irreductible"
author: "Marie Perrot-Dockès, Julien Chiquet"
date: "UseR2019"
output: 
  beamer_presentation :
    theme: "Rochester"
    colortheme: "vert"
header-includes:
- \usepackage{tikz}
- \usepackage{natbib}
- \usepackage{pgfplots}
- \newcommand{\Am}{\textrm{Argmin}}
- \newcommand{\bX}{\boldsymbol{X}}
- \newcommand{\bY}{\boldsymbol{Y}}
- \newcommand{\bS}{\boldsymbol{S}}
- \usetikzlibrary{arrows, decorations.markings}
- \usepackage{arydshln}
- \usepackage{url}
- \definecolor{outcome}{RGB}{165,77,106}
- \definecolor{dc}{RGB}{173,220,153}
- \definecolor{lt}{RGB}{153,173,220}
- \titlegraphic{\includegraphics[width=2.5cm]{logo_apt}\hspace*{4.75cm}~\includegraphics[width=2cm]{logo_inra}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "cairo_pdf", message = 'hide', warning = FALSE, cache = TRUE)
library(VariSel)
library(tidyverse)
library(extrafont)
library(ggsci)
require(tidymodels)
require(furrr)
require(rlang)
require(patchwork)
loadfonts()
 theme_set(theme_bw() + theme_bw() +
             theme(strip.background = element_rect(fill = "white"), 
                   text = element_text(face="bold", family="LM Roman 10", size=16) ))
 
scale_colour_discrete <- function(...)scale_color_uchicago()
scale_fill_discrete <- function(...)scale_fill_uchicago()
```

# A simple vision of the immune system :
\framesubtitle{Or how Astérix and Obélix can kick-off the Romans}
 

```{tikz, tikz-ex2, fig.cap = "DC Th dialogue", fig.ext = 'pdf', echo =FALSE}
\definecolor{vert}{RGB}{ 109,205,89}
\definecolor{bleu}{RGB}{93,139,172}
\definecolor{marron}{RGB}{172,126,93}
\definecolor{marronfonce}{RGB}{120,88,65}
\definecolor{i1}{RGB}{122,181,168}
\definecolor{i2}{RGB}{122,165,181}
\definecolor{i3}{RGB}{122,181,139}
\definecolor{dc}{RGB}{173,220,153}
\definecolor{dc2}{RGB}{123, 87, 229}
\definecolor{lt}{RGB}{153,173,220}
\definecolor{pert}{RGB}{220,153,173}
\tikzstyle{line} = [draw, -latex, color = marron, line width=1pt]
\tikzstyle{block} = [rectangle, draw, fill=vert!20, draw= vert!80, text = marronfonce,
text width=3.5em, text centered, rounded corners, minimum height=2em,line width=1pt]
\tikzstyle{block3} = [rectangle, draw, fill=vert!10, draw= vert!90, text = marronfonce,
text width=3em, text centered, rounded corners, minimum height=1.1em,line width=0.7pt]
\tikzstyle{inp} = [rectangle, draw,  text = marronfonce,
text width=4.4em, text centered, rounded corners, minimum height=1em,line width=1pt]
  \begin{tikzpicture}[node distance = 2cm, auto]
    % nodes
     \node(DC){\includegraphics[width= 0.22\textwidth]{cellule_dend}}; 
      \node[left of = DC, node distance = 3 cm,  color = pert](Pert){\includegraphics[width= 0.17\textwidth]{romain.png}}; 
       \node[right of = DC, node distance = 4 cm](TL){\includegraphics[width= 0.17\textwidth]{ast_ob}}; 
       \node[right of = TL, node distance = 3 cm](L){\includegraphics[width= 0.17\textwidth]{paf-asterix}};
        \node[below of = DC, node distance =2.2 cm,text width= 0.17\textwidth,text centered,color =dc!70!black]{ Dendritic cells};
 \node[below of = Pert, node distance =2.2 cm,text width= 0.22\textwidth,text centered,color =pert!70!black]{ Perturbators};
      \node[below of = TL, node distance =2.2 cm,text width= 0.22\textwidth, text centered,color =lt!70!black]{  \;\;\;\; Th \;\; \;\; lymphocytes};
      \node[above of = TL, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 1.2 cm, color =lt!90!black]{ Th responses};
      \node[above of =DC, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 2 cm, color =dc!90!black]{ DC signals};





       \path [line,color = dc] ([yshift=-0.8 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.8 cm]TL.west);
        \path [line,color = dc] ([yshift=-0.4 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.4 cm]TL.west);
         \path [line,color = dc] ([xshift=- 0.1 cm]DC.east) --  (TL.west);
         \path [line,color = dc] ([yshift=0.4 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.4 cm]TL.west);
          \path [line,color = dc] ([yshift=0.8 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.8 cm]TL.west);

\path [line, color = gray](Pert) -- (DC.west);
 \path [line,color = lt]([xshift= -0.4cm, yshift = -0.7cm]TL.east) -- ([yshift = -0.7cm]L.west);
  \path [line,color = lt]([xshift= -0.4cm, yshift = -0.35cm]TL.east) -- ([yshift = -0.35cm]L.west);
  \path [line,color = lt] ([xshift= -0.4cm]TL.east) -- (L.west);
  
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.35cm]TL.east) -- ([yshift = 0.35cm]L.west);
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.7cm]TL.east) -- ([yshift = 0.7cm]L.west);
        \end{tikzpicture}



```

\tiny{ Grandclaudon, M., Perrot-Dockès, M, Trichot, C et al. \textit{ A Quantitative Multivariate Model of Human Dendritic Cell-T Helper Cell Communication (March 15, 2019)}. Available at \url{http://dx.doi.org/10.2139/ssrn.3353217} }
<!--   \node[below of = DC, node distance =3.2 cm,text width= 0.17\textwidth,text centered,color =dc!70!black]{Sentinnelles};    -->
<!--  \node[below of = Pert, node distance =3.2 cm,text width= 0.22\textwidth,text centered,color =pert!70!black]{ Perturbation}; -->
<!--       \node[below of = TL, node distance =3.2 cm,text width= 0.22\textwidth, text centered,color =lt!70!black]{ Warriors}; -->
<!--       \node[below of = TL, node distance =3.2 cm,text width= 0.15\textwidth, text centered, xshift = 1.2 cm, color =lt!90!black]{Actions}; -->
<!--       \node[above of =DC, node distance =3.2 cm,text width= 0.15\textwidth, text centered, xshift = 2 cm, color =dc!90!black]{Warnings}; -->

# Experimental set up :
\framesubtitle{Ordralphabetix and Cétautomatix are two!}
 

```{tikz, tikz-ex2b, fig.cap = "DC Th dialogue", fig.ext = 'pdf', echo =FALSE, fig.num=1}
\definecolor{vert}{RGB}{ 109,205,89}
\definecolor{bleu}{RGB}{93,139,172}
\definecolor{marron}{RGB}{172,126,93}
\definecolor{marronfonce}{RGB}{120,88,65}
\definecolor{i1}{RGB}{122,181,168}
\definecolor{i2}{RGB}{122,165,181}
\definecolor{i3}{RGB}{122,181,139}
\definecolor{dc}{RGB}{173,220,153}
\definecolor{dc2}{RGB}{123, 87, 229}
\definecolor{lt}{RGB}{153,173,220}
\definecolor{pert}{RGB}{220,153,173}
\tikzstyle{line} = [draw, -latex, color = marron, line width=1pt]
\tikzstyle{block} = [rectangle, draw, fill=vert!20, draw= vert!80, text = marronfonce,
text width=3.5em, text centered, rounded corners, minimum height=2em,line width=1pt]
\tikzstyle{block3} = [rectangle, draw, fill=vert!10, draw= vert!90, text = marronfonce,
text width=3em, text centered, rounded corners, minimum height=1.1em,line width=0.7pt]
\tikzstyle{inp} = [rectangle, draw,  text = marronfonce,
text width=4.4em, text centered, rounded corners, minimum height=1em,line width=1pt]
  \begin{tikzpicture}[node distance = 2cm, auto]
      \node(DC){\includegraphics[width= 0.19\textwidth]{cellule_dend1}}; 
      \node[left of = DC, node distance = 3 cm,  color = pert](Pert){\includegraphics[width= 0.17\textwidth]{romain.png}}; 
       \node[right of = DC, node distance = 4 cm](TL){\includegraphics[width= 0.17\textwidth]{ast_ob}}; 
       \node[right of = TL, node distance = 3 cm](L){\includegraphics[width= 0.17\textwidth]{paf-asterix}};
        \node[below of = DC, node distance =2.2 cm,text width= 0.17\textwidth,text centered,color =dc!70!black]{ \textbf{Ordra}};
 \node[below of = Pert, node distance =2.2 cm,text width= 0.22\textwidth,text centered,color =pert!70!black]{ Perturbators};
      \node[below of = TL, node distance =2.2 cm,text width= 0.22\textwidth, text centered,color =lt!70!black]{  \;\;\;\; Th \;\; \;\; lymphocytes};
      \node[above of = TL, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 1.2 cm, color =lt!90!black]{ Th responses};
      \node[above of =DC, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 2 cm, color =dc!90!black]{ DC signals};





       \path [line,color = dc] ([yshift=-0.8 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.8 cm]TL.west);
        \path [line,color = dc] ([yshift=-0.4 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.4 cm]TL.west);
         \path [line,color = dc] ([xshift=- 0.1 cm]DC.east) --  (TL.west);
         \path [line,color = dc] ([yshift=0.4 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.4 cm]TL.west);
          \path [line,color = dc] ([yshift=0.8 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.8 cm]TL.west);

\path [line, color = gray](Pert) -- (DC.west);
 \path [line,color = lt]([xshift= -0.4cm, yshift = -0.7cm]TL.east) -- ([yshift = -0.7cm]L.west);
  \path [line,color = lt]([xshift= -0.4cm, yshift = -0.35cm]TL.east) -- ([yshift = -0.35cm]L.west);
  \path [line,color = lt] ([xshift= -0.4cm]TL.east) -- (L.west);
  
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.35cm]TL.east) -- ([yshift = 0.35cm]L.west);
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.7cm]TL.east) -- ([yshift = 0.7cm]L.west);
        \end{tikzpicture}

  \begin{tikzpicture}[node distance = 2cm, auto]
     \node(DC){\includegraphics[width= 0.19\textwidth]{cellule_dend2}}; 
      \node[left of = DC, node distance = 3 cm,  color = pert](Pert){\includegraphics[width= 0.17\textwidth]{romain.png}}; 
       \node[right of = DC, node distance = 4 cm](TL){\includegraphics[width= 0.17\textwidth]{ast_ob}}; 
       \node[right of = TL, node distance = 3 cm](L){\includegraphics[width= 0.17\textwidth]{paf-asterix}};
        \node[below of = DC, node distance =2.2 cm,text width= 0.17\textwidth,text centered,color =dc!70!black]{ \textbf{Cetau}};
 \node[below of = Pert, node distance =2.2 cm,text width= 0.22\textwidth,text centered,color =pert!70!black]{ Perturbators};
      \node[below of = TL, node distance =2.2 cm,text width= 0.22\textwidth, text centered,color =lt!70!black]{  \;\;\;\; Th \;\; \;\; lymphocytes};
      \node[above of = TL, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 1.2 cm, color =lt!90!black]{ Th responses};
      \node[above of =DC, node distance =1.5 cm,text width= 0.15\textwidth, text centered, xshift = 2 cm, color =dc!90!black]{ DC signals};





       \path [line,color = dc] ([yshift=-0.8 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.8 cm]TL.west);
        \path [line,color = dc] ([yshift=-0.4 cm,xshift=- 0.1 cm]DC.east) -- ([yshift=-0.4 cm]TL.west);
         \path [line,color = dc] ([xshift=- 0.1 cm]DC.east) --  (TL.west);
         \path [line,color = dc] ([yshift=0.4 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.4 cm]TL.west);
          \path [line,color = dc] ([yshift=0.8 cm, xshift=- 0.1 cm]DC.east) --  ([yshift=0.8 cm]TL.west);

\path [line, color = gray](Pert) -- (DC.west);
 \path [line,color = lt]([xshift= -0.4cm, yshift = -0.7cm]TL.east) -- ([yshift = -0.7cm]L.west);
  \path [line,color = lt]([xshift= -0.4cm, yshift = -0.35cm]TL.east) -- ([yshift = -0.35cm]L.west);
  \path [line,color = lt] ([xshift= -0.4cm]TL.east) -- (L.west);
  
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.35cm]TL.east) -- ([yshift = 0.35cm]L.west);
   \path [line,color = lt] ([xshift= -0.4cm, yshift = 0.7cm]TL.east) -- ([yshift = 0.7cm]L.west);
        \end{tikzpicture}

```


# Statistical modelling 
\framesubtitle{Understand the Gallic language}
 
\begin{itemize}
\item \textbf{Dataset description:}

\begin{itemize}
\item \textcolor{dc!70!black}{$\boldsymbol{X}$}:  $n\times p$ design matrix : the DC signals \includegraphics[width =0.12\textwidth]{cellule_dend}
\item \textcolor{lt!70!black}{$\boldsymbol{Y}$}:  $n\times q$  response matrix : the Th responses  \includegraphics[width =0.12\textwidth]{ast_ob}
\end{itemize}

\item \textbf{Question:} Which variables influence the responses?

\item \textbf{Approach:}
\begin{itemize}
 \item Variable selection in 
$$
\boldsymbol{Y}=\boldsymbol{XB}+\boldsymbol{E},
$$
where  \begin{itemize}
\item $\boldsymbol{B}$:   $p\times q$  \textbf{sparse} coefficients matrix
\item $\boldsymbol{E}$: $n\times q$ error matrix  with $$\forall  i\in\{1,\dots, n\}, \; (E_{i,1},\dots,E_{i,q})\stackrel{iid}{\sim}~\mathcal{N}(0,\boldsymbol{\Sigma}_q)$$

\end{itemize}
\item We take the dependence into account by estimating $\boldsymbol{\Sigma}_q$. 
\end{itemize}
\end{itemize}



# Statistical modelling 
\framesubtitle{Understand the Gallic language}
 
\begin{itemize}
\item \textbf{Dataset description:}

\begin{itemize}
\item \textcolor{dc!70!black}{$\boldsymbol{X}$}:  $n\times p$ design matrix : the DC signals \includegraphics[width =0.12\textwidth]{cellule_dend}
\item \textcolor{lt!70!black}{$\boldsymbol{Y}$}:  $n\times q$  response matrix : the Th responses  \includegraphics[width =0.12\textwidth]{ast_ob}
\end{itemize}

\item \textbf{Question:} Which variables influence the responses?

\item \textbf{Approach:}
\begin{itemize}
 \item Variable selection in 
$$
\boldsymbol{Y}=\boldsymbol{XB}+\boldsymbol{E},
$$
where  \begin{itemize}
\item $\boldsymbol{B}$:   $p\times q$  \textbf{sparse} coefficients matrix
\item $\boldsymbol{E}$: $n\times q$ error matrix  with $$\forall  i\in\{1,\dots, n\}, \; (E_{i,1},\dots,E_{i,q})\stackrel{iid}{\sim}~\mathcal{N}(0,\boldsymbol{\Sigma}_q)$$

\end{itemize}
\item We take the dependence into account by estimating $\boldsymbol{\Sigma}_q$. 
\end{itemize}
\end{itemize}


# Differents penalties : for different point of view

\begin{itemize}
\item \textcolor{outcome}{\textbf{Lasso}} : \textit{select variables without taking into account  potential links.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_L = \Am_{b} \left\lbrace||y-\bX b||_2^2 + \lambda ||b||_1 \right\rbrace,
\end{equation*}
\item \textcolor{outcome}{\textbf{Group-Lasso}}  : \textit{select a group of variables.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_G = \Am_{b_1, \dots, b_L} \left\lbrace||y-\sum_{1 \leq \ell \leq L}\bX_{(\ell)} b_{(\ell)}||_2^2 + \lambda\sum_{1 \leq \ell \leq L}\sqrt{p_\ell} ||b_\ell||_2\right\rbrace,
\end{equation*}
\item \textcolor{outcome}{\textbf{Fused-Lasso}} :  \textit{influence a group of variables to have the same coefficient.}
\begin{equation*}\label{fuselasso}
	\widehat{b}_F =\Am_{b}||y-\bX b||_2^2 + \left\lbrace\lambda_1\sum_{(i,j) \in \mathcal{G}} |b_i - b_j|+ \lambda_2 ||b||_1\right\rbrace,
\end{equation*}
\end{itemize}

# Differents penalties : for different point of view

\begin{itemize}
\item \textcolor{outcome}{\textbf{Lasso}} : \textit{select variables without taking into account  potential links.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_L = \Am_{b} \left\lbrace||y-\bX b||_2^2 + \lambda ||b||_1 \right\rbrace,
\end{equation*}
\item \textcolor{outcome}{\textbf{Group-Lasso}}  : \textit{select a group of variables.}
\begin{equation*} \label{grouplasso}
	\widehat{b}_G = \Am_{b_1, \dots, b_L} \left\lbrace||y-\sum_{1 \leq \ell \leq L}\bX_{(\ell)} b_{(\ell)}||_2^2 + \lambda\sum_{1 \leq \ell \leq L}\sqrt{p_\ell} ||b_\ell||_2\right\rbrace,
\end{equation*}
\item \textcolor{outcome}{\textbf{Fused-Lasso}} :  \textit{influence a group of variables to have the same coefficient.}
\begin{equation*}\label{fuselasso}
	\widehat{b}_F =\Am_{b}||y-\bX b||_2^2 + \left\lbrace\lambda_1\sum_{(i,j) \in \mathcal{G}} |b_i - b_j|+ \lambda_2 ||b||_1\right\rbrace,
\end{equation*}
\end{itemize}





<!-- # Statistical modelling  -->

<!-- \begin{table} -->
<!-- \centering -->
<!-- \begin{tabular}{l:c:rrr} -->
<!--   \hline -->
<!--  & \textcolor{outcome}{\textbf{dc}} & \multicolumn{3}{c}{\textcolor{outcome}{\textbf{DC\_sign}}} \\ -->
<!--   \hline -->
<!--  &  & IL12p70 & TNFa & IL10 \\ -->
<!--   \hline -->
<!-- 1 & Ordra & 32.11 & 859.27 & 10.1 \\ -->
<!--   2 & Ordra & 10.20 & 661.92 & 8.40 \\ -->
<!--   \vdots & \vdots & \vdots & \vdots & \vdots \\ -->
<!--   428 & Cétau & 9.90 & 711.26 & 22.54 \\ -->
<!--    \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->
<!-- \vspace{-0.5cm} -->
<!-- $$\Downarrow$$ -->

<!-- \vspace{-0.5cm} -->
<!-- \begin{table} -->
<!-- \centering -->
<!-- \begin{tabular}{l:rrrrrr} -->
<!--   \hline -->
<!-- & \multicolumn{6}{c}{\textcolor{outcome}{\textbf{X}}} \\ -->
<!--   \hline -->
<!--  & IL12p70 & TNFa & IL10  & IL12p70 & TNFa  & IL10  \\ -->
<!--   &  in Ordra & in Ordra & in Ordra  & in Cétau & in Cétau & in Cétau \\ -->
<!--   \hline -->
<!-- 1 &  32.11 & 859.27 & 10.1 &  0 & 0 & 0  \\ -->
<!--   2 & 10.20 & 661.92 & 8.40 &  0 & 0 & 0  \\ -->
<!--   \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots  \\ -->
<!--   428 &  0 & 0 & 0 &  9.90 & 711.26 & 22.54 \\ -->
<!--    \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->



#  VariSel for one model type 
```{r, include=FALSE}
set.seed(4)
load("raw_data.RData")
load("data_Tmod_sspdc.RData")
X <- Xbrt %>% as.data.frame() %>%
  mutate(pert_dc= pert_dc, dc= dc) %>%
  group_by(pert_dc, dc) %>%
  summarize_if(is.numeric, mean) %>% ungroup()  %>%
  as.data.frame()
# 
# Y <- Y_bxcx %>% as.data.frame() %>% 
#   mutate(pert_dc= pert_dc, dc= dc) %>% 
#   group_by(pert_dc, dc) %>% 
#   summarize_if(is.numeric, mean) %>% 
#   ungroup() %>% as.data.frame()
 DC_sign <- X_bxcx[,c(36,37, 40)]
 T_resp <- Y_bxcx[,c(2,10)]
# DC_sign <- X %>% select(-dc,-pert_dc) %>%
#   as.matrix() %>% scale()
# 
# T_resp <- Y %>% select(IFNg,  IL3) %>%
#   as.matrix() %>% scale()
# dc <- Y %>% pull(dc)
 dc <- as.character(dc)
 dc[dc == "bDC "] <- "Ordra"
dc[dc == "MoDC"] <- "Cetau"
```

```{r}
mod <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_regr") 
```


```{r, eval=FALSE}
X <-  model.matrix(~DC_sign:dc -1)
mod <-  train_VariSel( Y = T_resp, 
                       X = X, 
                       sepx = ":", 
                       type ="group_multi_regr")
```




<!-- Different parameters :  -->
<!-- \begin{itemize} -->
<!-- \item $\bY$ : Matrix containing the reponse values (here the Th signals for the different samples) -->
<!-- \item $\bX$ : optional design matrix of the model  -->
<!-- \item \textbf{type} : type of model -->
<!-- \item \textbf{regressors} : if X is not supplied the continuous values  -->
<!-- \item \textbf{group} : if the samples can be divided into groups  -->
<!-- \item \textbf{sepx} : if the columns of X can be divided into group -->
<!-- \item \textbf{type\_S12\_inv} : specified the method use to estimate $\bS$ -->
<!-- \end{itemize} -->



#  VariSel for one model type 



```{r}
mod <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_regr") 
```


```{r, eval=FALSE}
X <-  model.matrix(~DC_sign:dc -1)
mod <-  train_VariSel( Y = T_resp, 
                       X = X, 
                       sepx = ":", 
                       type ="group_multi_regr")
```

#  VariSel for one model type : Outcome 

```{r}
plot(mod)
```

# Different modelling strategy 



```{r, echo=FALSE}
col <- pal_uchicago()(6)
mod <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_regr") 
m2 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_both")
m3 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="fused_multi_both")
m4 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="fused_multi_regr")
m5 <-   train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="lasso_multi")

```

```{r}
compar_path(mods = list(mod,m2,m3,m4))
```

# Different modelling strategy 

```{r, echo=FALSE}
col <- pal_uchicago()(6)
mod <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_regr") 
m2 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="group_multi_both")
m3 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="fused_multi_both")
m4 <-  train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="fused_multi_regr")
m5 <-   train_VariSel( Y = T_resp,
                       regressors = DC_sign, 
                       group = dc, 
                       type ="lasso_multi")

```

```{r}
compar_path(mods = list(mod,m2,m3,m4))
```

# Models selection 

```{r}
ct <- compar_type( Y = T_resp, regressors = DC_sign,  
                   group = dc, 
 types = c("group_multi_regr" , "group_multi_both" ,
  "fused_multi_regr", "fused_multi_both",
  "lasso_multi" ), times = 10)

```

```{r, echo=FALSE, fig.width=6,fig.height=3, message=FALSE}
plot_ct(ct) + labs(title ='')
```




# Best models representation

```{r, fig.height =5.5}
bm <- get_best_models(ct,criterion = "MSE_boot")
plot_md(bm)
```


# Conclusion

This is an R package to perform variable selection in multivariate linear models. 
It can
\begin{itemize}
\item Associate explicative variables 
\item Associate responses 
\item Associate both explicative  variables and responses 
\item Let all variables 'free', without assoiciating any of them
\end{itemize}


Come and see the vignette!
\textcolor{outcome}{https://github.com/Marie-PerrotDockes/VariSel}

